---
title: "Look at correlations between FD and MLD."
output: html_notebook
---

Load up the usual packages.
```{r}
pacman::p_load(tidyr, dplyr, ggplot2, plyr, nlme, reshape2, gridExtra, grid)
```

Get the data.
```{r}
raw_data <- read.csv2("C:/Users/csuehs/Dropbox/SCANN'AIR_fractal_dimension/FD_data/Expi.csv")
raw_data2 <- read.csv("C:/Users/csuehs/Dropbox/SCANN'AIR_princeps/data-verification/Data_scannair_NA_corrections_isaetcarey_20170427.csv")
```

Extract the appropriate variables. We are interested in raw FD values for attentuation at -850HU and according to two factors: 

* inspiratory versus expiratory measures;
* upper versus lower lung.
      ---slice 1
      ---slice 10
```{r}

df_fd <- data.frame(raw_data$numpat,
                  raw_data$I.850.H1,
                  raw_data$I.850.H10,
                  raw_data$I.850.L1,
                  raw_data$I.850.L10,
                  raw_data$E.850.H1,
                  raw_data$E.850.H10,
                  raw_data$E.850.L1,
                  raw_data$E.850.L10)

colnames(df_fd) <- c("patient",
                   "fd_i_up_1", 
                    "fd_i_up_10", 
                    "fd_i_lo_1", 
                    "fd_i_lo_10", 
                    "fd_e_up_1", 
                    "fd_e_up_10", 
                    "fd_e_lo_1", 
                    "fd_e_lo_10")



df_fd$fd_ei_up_1 <- df_fd$fd_e_up_1 / df_fd$fd_i_up_1
df_fd$fd_ei_up_10 <- df_fd$fd_e_up_10 / df_fd$fd_i_up_10

df_fd$fd_ei_lo_1 <- df_fd$fd_e_lo_1 / df_fd$fd_i_lo_1
df_fd$fd_ei_lo_10 <- df_fd$fd_e_lo_10 / df_fd$fd_i_lo_10


df_fd$fd_i_up <- (df_fd$fd_i_up_1 + df_fd$fd_i_up_10)/2
df_fd$fd_i_lo <- (df_fd$fd_i_lo_1 + df_fd$fd_i_lo_10)/2

df_fd$fd_e_up <- (df_fd$fd_e_up_1 + df_fd$fd_e_up_10)/2
df_fd$fd_e_lo <- (df_fd$fd_e_lo_1 + df_fd$fd_e_lo_10)/2

df_fd$fd_ei_up <- (df_fd$fd_ei_up_1 + df_fd$fd_ei_up_10)/2
df_fd$fd_ei_lo <- (df_fd$fd_ei_lo_1 + df_fd$fd_ei_lo_10)/2

```

Extract the appropriate MLD variables.
```{r}
df_mld <- data.frame(raw_data2$numpat,
                     raw_data2$I_MLD_H_1,
                     raw_data2$I_MLD_B_1,
                     raw_data2$E_MLD_H_1,
                     raw_data2$E_MLD_B_1)
    

colnames(df_mld) <- c("patient",
                      "mld_i_up",
                      "mld_i_lo",
                      "mld_e_up",
                      "mld_e_lo")

df_mld$mld_ei_up <- df_mld$mld_e_up/df_mld$mld_i_up
df_mld$mld_ei_lo <- df_mld$mld_e_lo/df_mld$mld_i_lo

```

Do some datawrangling in preparation for a ggplots.

* grouping column: inspiratory vs expiratory
* grouping column: upper vs lower
* value column: FD 
* value column: MLD
```{r}

df <- merge(df_fd, df_mld, by = "patient")

n <- c("breath", "region", "fd", "mld")

temp_i_up <- data.frame(rep("inspiratory", dim(df)[1]),
                    rep("upper", dim(df)[1]),
                    df$fd_i_up,
                    df$mld_i_up)

temp_i_lo <- data.frame(rep("inspiratory", dim(df)[1]),
                    rep("lower", dim(df)[1]),
                    df$fd_i_lo,
                    df$mld_i_lo)

temp_e_up <- data.frame(rep("expiratory", dim(df)[1]),
                    rep("upper", dim(df)[1]),
                    df$fd_e_up,
                    df$mld_e_up)

temp_e_lo <- data.frame(rep("expiratory", dim(df)[1]),
                    rep("lower", dim(df)[1]),
                    df$fd_e_lo,
                    df$mld_e_lo)


colnames(temp_i_up) <- n
colnames(temp_i_lo) <- n
colnames(temp_e_up) <- n
colnames(temp_e_lo) <- n



long <- rbind(temp_i_up,
              temp_i_lo,
              temp_e_up,
              temp_e_lo)

rm(temp_i_up, temp_i_lo, temp_e_up, temp_e_lo, n)

```

Do some datawrangling for e/i.

* grouping variable: upper versus lower
* value variable: fd
* value variable: md

```{r}

temp_ei_up <- data.frame(rep("upper", dim(df)[1]),
                        df$fd_ei_up,
                        df$mld_ei_up)

temp_ei_lo <- data.frame(rep("lower", dim(df)[1]),
                        df$fd_ei_lo,
                        df$mld_ei_lo)

n <- c("region", "fd", "mld")

colnames(temp_ei_up) <- n
colnames(temp_ei_lo) <- n

long_ei <- rbind(temp_ei_up,
                 temp_ei_lo)


```



Get spearman correlation coefficients and p values.
```{r}
r_i_up <- round(cor(df$fd_i_up, df$mld_i_up, method = c("pearson")), digits=2)
p_i_up <- round(cor.test(df$fd_i_up, df$mld_i_up, method = c("pearson"))[[3]], digits = 3)


r_e_up <- round(cor(df$fd_e_up, df$mld_e_up, method = c("pearson")), digits=2)
p_e_up <- round(cor.test(df$fd_e_up, df$mld_e_up, method = c("pearson"))[[3]], digits = 3)


r_i_lo <- round(cor(df$fd_i_lo, df$mld_i_lo, method = c("pearson")), digits=2)
p_i_lo <- round(cor.test(df$fd_i_lo, df$mld_i_lo, method = c("pearson"))[[3]], digits = 3)


r_e_lo <- round(cor(df$fd_e_lo, df$mld_e_lo, method = c("pearson")), digits=2)
p_e_lo <- round(cor.test(df$fd_e_lo, df$mld_e_lo, method = c("pearson"))[[3]], digits = 3)

r_ei_up <- round(cor(df$fd_ei_up, df$mld_ei_up, method = c("pearson")), digits=2)
p_ei_up <- round(cor.test(df$fd_ei_up, df$mld_ei_up, method = c("pearson"))[[3]], digits = 3)

r_ei_lo <- round(cor(df$fd_ei_lo, df$mld_ei_lo, method = c("pearson")), digits=2)
p_ei_lo <- round(cor.test(df$fd_ei_lo, df$mld_ei_lo, method = c("pearson"))[[3]], digits = 3)

label_i_up <- paste("R^2 ==", r_i_up,"* '; P < 0.0001'")
label_e_up <- paste("R^2 ==", r_e_up,"* '; P < 0.0001'")

label_i_lo <- paste("R^2 ==", r_i_lo,"* '; P = 0.003'")
label_e_lo <- paste("R^2 ==", r_e_lo,"* '; P < 0.0001'")

label_ei_up <- paste("R^2 ==", r_ei_up,"* '; P < 0.0001'")
label_ei_lo <- paste("R^2 ==", r_ei_up,"* '; P < 0.0001'")
```



Make inspiratory and expiratory subsets
```{r}

long_insp <- long[long$breath == "inspiratory",]

long_exp <- long[long$breath == "expiratory",]
```




Make a ggplot for inspiratory, upper.
```{r}


plot_i_up <- ggplot(long_insp[long_insp$region == "upper",], aes(x = fd, y = mld))+
    geom_point()+
    geom_smooth(method=lm)+
    ylab(expression(paste(MLD[I]," (-HU)")))+
    xlab(expression(paste(FD[I-850])))+
    annotate("text", x = 1.6, y = 850, label=label_i_up, parse = TRUE, hjust=0)+
    ggtitle("Inspiratory")+
    theme(plot.title = element_text(hjust = 0.5))

plot_i_up

```
Make a ggplot for inspiratory, lower.
```{r}


plot_i_lo <- ggplot(long_insp[long_insp$region == "lower",], aes(x = fd, y = mld))+
    geom_point()+
    geom_smooth(method=lm)+
    ylab(expression(paste(MLD[I]," (-HU)")))+
    xlab(expression(paste(FD[I-850])))+
    annotate("text", x = 1.6, y = 850, label=label_i_lo, parse = TRUE, hjust=0)+
    ggtitle("   ")+
    theme(plot.title = element_text(hjust = 0.5))

plot_i_lo

```

Make a ggplot for expiratory, upper.
```{r}


plot_e_up <- ggplot(long_exp[long_exp$region == "upper",], aes(x = fd, y = mld))+
    geom_point()+
    geom_smooth(method=lm)+
    ylab(expression(paste(MLD[E]," (-HU)")))+
    xlab(expression(paste(FD[E-850])))+
    annotate("text", x = 0.1, y = 730, label=label_e_up, parse = TRUE, hjust=0)+
    ggtitle("Expiratory")+
    theme(plot.title = element_text(hjust = 0.5))

plot_e_up

```

Make a ggplot for expiratory, lower.
```{r}


plot_e_lo <- ggplot(long_exp[long_exp$region == "lower",], aes(x = fd, y = mld))+
    geom_point()+
    geom_smooth(method=lm)+
    ylab(expression(paste(MLD[E]," (-HU)")))+
    xlab(expression(paste(FD[E-850])))+
    annotate("text", x = 0.2, y = 700, label=label_e_lo, parse = TRUE, hjust=0)+
    ggtitle("   ")+
    theme(plot.title = element_text(hjust = 0.5))

plot_e_lo

```



Make a ggplot for ei values, upper.

```{r}

plot_ei_up <- ggplot(long_ei[long_ei$region == "upper",], aes(x = fd, y = mld))+
    geom_point()+
    geom_smooth(method=lm)+
    ylab(expression(MLD[E/I]))+
    xlab(expression(paste(FD[E/I-850])))+
    annotate("text", x = 0.05, y = 0.9, label=label_ei_up, parse = TRUE, hjust=0)+
    ggtitle("E/I")+
    theme(plot.title = element_text(hjust = 0.5))

plot_ei_up

```
Make a ggplot for ei values, upper.

```{r}

plot_ei_lo <- ggplot(long_ei[long_ei$region == "lower",], aes(x = fd, y = mld))+
    geom_point()+
    geom_smooth(method=lm)+
    ylab(expression(MLD[E/I]))+
    xlab(expression(paste(FD[E/I-850])))+
    annotate("text", x = 0.05, y = 0.88, label=label_ei_lo, parse = TRUE, hjust=0)+
    ggtitle("   ")+
    theme(plot.title = element_text(hjust = 0.5))

plot_ei_lo

```


Combine everything into 1 graphic.

```{r}


fff <- list(plot_i_up, 
             plot_e_up,
             plot_ei_up,
             plot_i_lo, 
             plot_e_lo, 
             plot_ei_lo)

ddd <- tableGrob(c("", "Upper \nlung", "Lower \nlung"), theme= ttheme_minimal())

rg <- arrangeGrob(grobs = fff, ncol=3, top = textGrob("",gp=gpar(fontsize=18)))
grid.newpage()
grid.draw(cbind(ddd, rg, size = "last"))


```

Print to the appropriate file.
```{r}
pdf (file = "C:/Users/csuehs/Dropbox/SCANN'AIR_fractal_dimension/Submission2_Investigative_Radiology/figure_3/figure_3_correlation_plots.pdf", width = 12, height=6)



grid.newpage()
grid.draw(cbind(ddd, rg, size = "last"))


dev.off()
```






